---
title: "R Notebook"
output: html_notebook
---

# Load in libraries

```{r}

library(tidyverse)
library(janitor)
library(lubridate)

```

# Read in data

```{r}

flights <- read_csv("clean_data/flights_clean.csv")
airlines <- read_csv("raw_data/airlines.csv")
airports <- read_csv("raw_data/airports.csv")
weather <- read_csv("raw_data/weather.csv")
nyc_daily_weather <- read_csv("raw_data/daily_weather_data_2017.csv")

```

```{r}



```

```{r}
# daily variance in number of delayed flights - Newark
flights %>% 
  filter(!is.na(is_delayed)) %>% 
  filter(origin == "EWR") %>% 
  # group_by(sched_dep_hour) %>% 
  # mutate(props_delayed) 
  # filter(dep_delay > 0) %>% 
  ggplot(aes(x = hour, fill = is_delayed)) +
  geom_histogram(col = "white", binwidth = 1) +
  facet_wrap(~is_delayed, nrow = 3)

# daily variance in number of delayed flights - JFK
flights %>% 
  filter(!is.na(is_delayed)) %>% 
  filter(origin == "JFK") %>% 
  # group_by(sched_dep_hour) %>% 
  # mutate(props_delayed) 
  # filter(dep_delay > 0) %>% 
  ggplot(aes(x = hour, fill = is_delayed)) +
  geom_histogram(col = "white", binwidth = 1) +
  facet_wrap(~is_delayed, nrow = 3)

# daily variance in number of delayed flights - LGA
flights %>% 
  filter(!is.na(is_delayed)) %>% 
  filter(origin == "LGA") %>% 
  # group_by(sched_dep_hour) %>% 
  # mutate(props_delayed) 
  # filter(dep_delay > 0) %>% 
  ggplot(aes(x = hour, fill = is_delayed)) +
  geom_histogram(col = "white", binwidth = 1) +
  facet_wrap(~is_delayed, nrow = 3)

```

```{r}
# proportion of flights delayed throughout the day
flights %>%
  filter(!is.na(is_delayed)) %>% 
  group_by(origin, hour) %>%
  summarise(n_delayed = sum(is_delayed),
            props_delayed = n_delayed / n()) %>% 
  ggplot(aes(x = hour, y = props_delayed, fill = origin)) +
  geom_col(col = "white", binwidth = 1) +
  facet_wrap(~origin, ncol = 1) +
  scale_x_continuous(breaks = 0:23)

# average delay in mins throughout the day
flights %>%
  filter(is_delayed == 1) %>% 
  group_by(origin, hour) %>%
  summarise(avg_delay = mean(dep_delay)) %>% 
  ggplot(aes(x = hour, y = avg_delay, fill = origin)) +
  geom_col(col = "white", binwidth = 1) +
  facet_wrap(~origin, ncol = 1) +
  scale_x_continuous(breaks = 0:23)


flights %>%
  filter(is.na(is_delayed))

flights %>% 
  ggplot(aes(x = hour)) +
  geom_histogram(binwidth = 1, col = "white") +
  scale_x_continuous(breaks = 0:23)

```

# Flight cancellations

```{r}
# cancellations over the year
seasonal_cancellations <- flights %>% 
  group_by(origin, month) %>% 
  summarise(cancellations = sum(is_cancelled),
            prop_cancelled = cancellations / n())

# number of cancellations over the year
seasonal_cancellations %>% 
  ggplot(aes(x = month, y = cancellations, col = origin)) +
  geom_point() +
  geom_line() +
  scale_x_continuous(breaks = 1:24) +
  labs(title = "Number of cancellations across the year")

# proportion of cancellations over the year
seasonal_cancellations %>% 
  ggplot(aes(x = month, y = prop_cancelled, col = origin)) +
  geom_point() +
  geom_line() +
  scale_x_continuous(breaks = 1:24) +
  labs(title = "Proportion of cancellations across the year")

```


```{r}
# cancellations over the day
diurnal_cancellations <- flights %>% 
  group_by(origin, hour) %>% 
  summarise(cancellations = sum(is_cancelled),
            prop_cancelled = cancellations / n())

# number of cancellations over the day
diurnal_cancellations %>% 
  ggplot(aes(x = hour, y = cancellations, col = origin)) +
  geom_point() +
  geom_line() +
  scale_x_continuous(breaks = 1:24) +
  labs(title = "Number of cancellations across the day")

# proportion of cancellations over the day
diurnal_cancellations %>% 
  ggplot(aes(x = hour, y = prop_cancelled, col = origin)) +
  geom_point() +
  geom_line() +
  scale_x_continuous(breaks = 1:24) +
  labs(title = "Proportion of cancellations across the day")

```

# Delays and cancellations per airline

```{r}
# total number of flights scheduled per airline - Newark
flights %>% 
  filter(origin == "EWR") %>% 
  group_by(carrier) %>% 
  summarise(total_scheduled_flights = n()) %>% 
  ggplot(aes(x = reorder(carrier, - total_scheduled_flights),
             y = total_scheduled_flights)) +
  geom_col() +
  labs(title = "Scheduled Annual Flights per airline - Newark Airport",
       x = "Airline",
       y = "Number of Flights")

# total number of flights scheduled per airline - JFK
flights %>% 
  filter(origin == "JFK") %>% 
  group_by(carrier) %>% 
  summarise(total_scheduled_flights = n()) %>% 
  ggplot(aes(x = reorder(carrier, - total_scheduled_flights),
             y = total_scheduled_flights)) +
  geom_col() +
  labs(title = "Scheduled Annual Flights per airline - JFK Airport",
       x = "Airline",
       y = "Number of Flights")

# total number of flights scheduled per airline - LGA
flights %>% 
  filter(origin == "LGA") %>% 
  group_by(carrier) %>% 
  summarise(total_scheduled_flights = n()) %>% 
  ggplot(aes(x = reorder(carrier, - total_scheduled_flights),
             y = total_scheduled_flights)) +
  geom_col() +
  labs(title = "Scheduled Annual Flights per airline - LGA Airport",
       x = "Airline",
       y = "Number of Flights")

```


```{r}
airline_delays_cancels <- flights %>% 
  group_by(carrier) %>% 
  summarise(total_flights = n(),
            delayed_flights = sum(is_delayed, na.rm = TRUE),
            prop_delayed_flights = mean(is_delayed, na.rm = TRUE),
            cancelled_flights = sum(is_cancelled),
            prop_cancelled_flights = mean(is_cancelled))

# airline_delays_cancels %>% 
#   ggplot(aes(x = carrier, y = delayed_flights, fill = total_flights)) +
#   geom_col()

airline_delays_cancels %>% 
  ggplot(aes(x = carrier, y = prop_delayed_flights, fill = total_flights)) +
  geom_col() +
  labs(title = "Proportion of flights delayed per airline") +
  scale_y_continuous(labels = scales::percent)

# airline_delays_cancels %>% 
#   ggplot(aes(x = carrier, y = cancelled_flights, fill = total_flights)) +
#   geom_col()

airline_delays_cancels %>% 
  ggplot(aes(x = carrier, y = prop_cancelled_flights, fill = total_flights)) +
  geom_col() +
  labs(title = "Proportion of flights cancelled per airline") +
  scale_y_continuous(labels = scales::percent)

```

```{r}

# total number of flights scheduled per airline - Newark with delays and cancellations
delays_cancellations <- flights %>%
  group_by(origin, carrier) %>% 
  summarise(delayed_flights = sum(is_delayed, na.rm = TRUE),
            cancelled_flights = sum(is_cancelled),
            ontime_flights = n() - delayed_flights - cancelled_flights) %>%
  pivot_longer(ends_with("flights"), names_to = "flight_status",
               values_to = "num_flights") %>% 
  mutate(flight_status = str_remove(flight_status, "_flights"),
         flight_status = factor(flight_status,levels = c("ontime", "delayed",
                                                         "cancelled")))

delays_cancellations %>% 
  filter(origin == "EWR") %>% 
  ggplot(aes(x = reorder(carrier, - num_flights), y = num_flights,
             fill = flight_status)) +
  geom_col(col = "black") +
  labs(title = "Scheduled Annual Flights per airline - Newark Airport",
       x = "Airline",
       y = "Number of Flights")

delays_cancellations %>% 
  filter(origin == "JFK") %>% 
  ggplot(aes(x = reorder(carrier, - num_flights), y = num_flights,
             fill = flight_status)) +
  geom_col(col = "black") +
  labs(title = "Scheduled Annual Flights per airline - JFK Airport",
       x = "Airline",
       y = "Number of Flights")

delays_cancellations %>% 
  filter(origin == "LGA") %>% 
  ggplot(aes(x = reorder(carrier, - num_flights), y = num_flights,
             fill = flight_status)) +
  geom_col(col = "black") +
  labs(title = "Scheduled Annual Flights per airline - La Guardia Airport",
       x = "Airline",
       y = "Number of Flights")


```



```{r}

weather_daily <- read_csv("clean_data/daily_weather.csv")

```

# Dealing with missing values

It is assumed that the weather between the 3 NY airports are similar so any missing values for:
- wind_dir (wind direction)
- wind_speed 
- wind_gust 
- visib

Will be replaced by the average of the other airports data for that date.

```{r}

weather <- read_csv("clean_data/weather_clean.csv")

# Impute missing wind_dir with average values from other airports
hourly_weather_clean <- weather %>% 
  left_join(avg_weather, "time_hour") %>% 
  mutate(wind_dir = if_else(wind_dir == 0, NA_real_, wind_dir),
         wind_speed = if_else(wind_speed == 0, NA_real_, wind_speed),
         wind_gust = if_else(wind_gust == 0, NA_real_, wind_gust)) %>% 
  mutate(wind_dir = coalesce(wind_dir, avg_wind_dir),
         wind_speed = coalesce(wind_speed, avg_wind_speed),
         wind_gust = coalesce(wind_gust, avg_wind_gust),
         visib = coalesce(visib, avg_visib)) %>% 
  select(time_hour, origin, year, month, day, hour, wind_dir, wind_speed,
         wind_gust, visib) %>% 
  drop_na()

daily_weather_data <- hourly_weather_clean %>%
  mutate(date = as_date(time_hour), .after = time_hour) %>% 
  group_by(date, origin) %>% 
  summarise(avg_wind_speed = mean(wind_speed),
            max_wind_speed = max(wind_gust),
            avg_visib = mean(visib)) %>%
  arrange(origin, date) %>% 
  left_join(weather_daily, by = c("date", "origin"))
  
daily_weather_data

```

```{r}

# avg and max wind speeds
daily_weather_data %>% 
  filter(origin == "EWR") %>% 
  ggplot(aes(x = date)) +
  geom_line(aes(y = avg_wind_speed), col = "blue") +
  geom_line(aes(y = max_wind_speed), col = "red")

daily_weather_data %>% 
  filter(origin == "EWR") %>% 
  ggplot(aes(x = avg_wind_speed)) +
  geom_histogram(binwidth = 1) # normal distribution with positive skew

# avg visibility
daily_weather_data %>% 
  filter(origin == "EWR") %>% 
  ggplot(aes(x = date)) +
  geom_line(aes(y = avg_visib), col = "blue")

daily_weather_data %>% 
  filter(origin == "EWR") %>% 
  ggplot(aes(x = avg_visib)) +
  geom_histogram() # predominantly 10 (max value), long tail to the left

# avg_precip
daily_weather_data %>% 
  filter(origin == "EWR") %>% 
  ggplot(aes(x = date)) +
  geom_line(aes(y = prcp), col = "blue")

daily_weather_data %>% 
  filter(origin == "EWR") %>% 
  ggplot(aes(x = prcp)) +
  geom_histogram() # predominantly 0 (min value), long tail to the right

# snow and snow depths
daily_weather_data %>% 
  filter(origin == "EWR") %>% 
  ggplot(aes(x = date)) +
  geom_line(aes(y = snow), col = "blue") +
  geom_line(aes(y = snow_depth), col = "red")

# there's always a lag between snowfall and snow depth indicating the build up 
# of snow and subsequent melting of snow

daily_weather_data %>% 
  filter(origin == "EWR") %>% 
  ggplot(aes(x = snow)) +
  geom_histogram(binwidth = 1) # mostly 0 values, skewed to the right.

# min, avg and max temperature
daily_weather_data %>% 
  filter(origin == "EWR") %>% 
  ggplot(aes(x = date)) +
  geom_line(aes(y = tavg), col = "green") +
  geom_line(aes(y = tmax), col = "red") +
  geom_line(aes(y = tmin), col = "blue")

```


# Analysis of Weather Data - Newark Airport

```{r}

# wind_speed
hourly_weather_clean %>% 
  filter(origin == "EWR") %>%
  ggplot(aes(x = time_hour)) +
  geom_line(aes(y = wind_speed), col = "red") +
  geom_line(aes(y = wind_gust), col = "blue")

# wind_gust
hourly_weather_clean %>% 
  filter(origin == "EWR") %>%
  ggplot(aes(x = time_hour, y = visib)) +
  geom_line()

```

```{r}

flights %>% 
  mutate(delayed = if_else(dep_delay > 0, TRUE, FALSE), .after = dep_delay)

flights %>% 
  ggplot(aes(x = dep_delay, y = arr_delay, colour = carrier)) +
  geom_point() +
  facet_wrap(~carrier)

```

# Summary Statistics for flights and delays

```{r}

# flights <- flights %>% 
#   mutate(delayed = if_else(dep_delay > 0, TRUE, FALSE), .after = dep_delay)
# 
# newark_flights <- flights %>% 
#   filter(origin == "EWR")

```

## Number of monthly flights 

```{r}

flights %>% 
  group_by(origin, month) %>% 
  summarise(count = n()) %>% 
  ggplot(aes(x = month, y = count, colour = origin)) +
  geom_line() +
  scale_x_continuous(breaks = 1:12)

```

Newark has the highest number of flights in NY with approximately 10,000 per
month between March and August. There is a peak in August.

## Proportion of flights which are delayed across 3 NY airports.

```{r}

flights %>% 
  group_by(origin) %>% 
  summarise(n = n(), prop_delayed = mean(is_delayed, na.rm = TRUE))

```

Newark Airport flights are delayed 38.6% of the time whereas JFK has only 37.3%
delayed flights and La Guardia has 35.4% delayed flights. Newark has the highest
proportion of delayed flights of the 3 airports compared here but the most 
flights. So why?

## Which routes are delayed the most?

```{r}

flights %>%
  group_by(origin, dest) %>% 
  summarise(prop_delayed = mean(is_delayed, na.rm = TRUE)) %>% 
  pivot_wider(names_from = origin, values_from = prop_delayed) %>% 
  arrange(desc(EWR)) %>% 
  left_join(airports, by = c("dest" = "faa")) %>% 
  select(dest, name, EWR, JFK, LGA)

```

## Average Delay Length 

```{r}

flights %>% 
  group_by(origin) %>% 
  filter(is_delayed == 1) %>% 
  summarise(avg_dep_delay = mean(dep_delay),
            med_dep_delay = median(dep_delay))

```

Average departure delay at Newark is approximately 48.6mins compared to 48.6mins
at JFK (the same) and 50.7mins at LGA. However, the median delay time is quite a
bit higher at Newark at 26mins, compared to 22 and 23 for JFK and LGA, 
respectively.

## Average monthly delay time

```{r}

monthly_delays <- flights %>% 
  group_by(origin, month) %>% 
  filter(is_delayed == 1) %>% 
  summarise(avg_dep_delay = mean(dep_delay),
            med_dep_delay = median(dep_delay),
            count = n())

monthly_delays

monthly_delays %>% 
  ggplot(aes(x = month, y = avg_dep_delay, col = origin)) +
  geom_line() +
  scale_x_continuous(breaks = 1:12) +
  labs(title = "Average monthly delay duration",
       x = "Month",
       y = "Delay Duration (mins)")

# monthly_delays %>% 
#   ggplot(aes(x = month, y = med_dep_delay, col = origin)) +
#   geom_line() +
#   scale_x_continuous(breaks = 1:12)

monthly_delays %>% 
  ggplot(aes(x = month, y = count, col = origin)) +
  geom_line() +
  scale_x_continuous(breaks = 1:12) +
  labs(title = "Monthly flight delays",
       x = "Month",
       y = "Delays")

```

The average monthly delay duration peaks in the summer in July, showing a strong
seasonal variation. However, the number of flight delays stays high from March 
to August (~4000 delayed flights per month). This tells us that despite July not
being the month with the most delays, there must be something else which is causing
the delays to be on average, longer in this month. For example, there is a
significant spike in the number of delays in December and January however the
average delay durations are at an annual low (around 43mins) compared to the
peak in July of 62 mins.

## Are there any planes which are delayed more frequently?

```{r}

flights %>% 
  group_by(tailnum) %>% 
  summarise(prop_delayed = mean(is_delayed, na.rm = TRUE),
            n = n()) %>% 
  arrange(desc(n), desc(prop_delayed)) %>% 
  ggplot(aes(y = n, x = prop_delayed)) +
  geom_point()


```

```{r}

flights %>% 
  left_join(planes, "tailnum") %>% 
  group_by(seats) %>% 
  summarise(prop_delayed = mean(is_delayed, na.rm = TRUE)) %>% 
  ggplot(aes(x = seats, y = prop_delayed)) +
  geom_point() +
  geom_smooth(method = "lm")

```


Runways:
Newark: 290 & 40
JFK & LGA: 310 & 40

```{r}
# newark runways and wind map
hourly_weather_clean %>% 
  filter(origin == "EWR") %>% 
  ggplot(aes(x = wind_dir)) +
  geom_histogram(binwidth = 10) +
  coord_polar() +
  scale_x_continuous(n.breaks = 12) +
  # add runway 4 (40 degrees)
  geom_vline(xintercept = 40, col = "blue") + 
  geom_vline(xintercept = 220, col = "blue") +
  # add Newark runway 29 (290 degrees)
  geom_vline(xintercept = 290, col = "green") +
  geom_vline(xintercept = 110, col = "green") +
  labs(title = "Newark Runways and Wind Sock")

# JFK runway and wind map
hourly_weather_clean %>% 
  filter(origin == "JFK") %>% 
  ggplot(aes(x = wind_dir)) +
  geom_histogram(binwidth = 10) +
  coord_polar() +
  scale_x_continuous(n.breaks = 12) +
  # add runway 4 (40 degrees)
  geom_vline(xintercept = 40, col = "blue") + 
  geom_vline(xintercept = 220, col = "blue") +
  # add JFK & LGA runway 31 (310 degrees)
  geom_vline(xintercept = 310, col = "red") +
  geom_vline(xintercept = 130, col = "red") +
  labs(title = "JFK Runways and Wind Sock")

# LGA runway and wind map
hourly_weather_clean %>% 
  filter(origin == "LGA") %>% 
  ggplot(aes(x = wind_dir)) +
  geom_histogram(binwidth = 10) +
  coord_polar() +
  scale_x_continuous(n.breaks = 12) +
  # add runway 4 (40 degrees)
  geom_vline(xintercept = 40, col = "blue") + 
  geom_vline(xintercept = 220, col = "blue") +
  # add JFK & LGA runway 31 (310 degrees)
  geom_vline(xintercept = 310, col = "red") +
  geom_vline(xintercept = 130, col = "red") +
  labs(title = "LGA Runways and Wind Sock")


avg_wind_dir <- hourly_weather_clean %>%
  group_by(origin) %>% 
  mutate(wind_dir = if_else(wind_dir < 180, wind_dir + 360, wind_dir)) %>% 
  summarise(avg_wind_dir = mean(wind_dir)) 

avg_wind_dir


```

Appears to be spikes in the frequency of wind direction for 310 and 40 (which
matches well with the runway angles). However, 2017 saw quite a large proportion
of wind directions heading 335 degrees. Newark runways have the biggest
difference between runway angle and wind_dir which may be a reason for the
number of delays at Newark.

```{r}

hourly_weather_clean %>% 
  # full_join(weather_clean, "time_hour") %>%
  # select(origin.x, wind_dir.x, wind_dir.y, time_hour) %>% 
  filter(origin == "JFK") %>% 
  #filter(!is.na(wind_dir.x)) %>% 
  ggplot(aes(x = wind_dir)) +
  geom_histogram(binwidth = 10) +
  coord_polar() +
    # add runway 4 (40 degrees)
  geom_vline(xintercept = 40, col = "blue") + 
  geom_vline(xintercept = 220, col = "blue") +
  # add JFK & LGA runway 31 (310 degrees)
  geom_vline(xintercept = 310, col = "red") +
  geom_vline(xintercept = 130, col = "red")

weather %>% 
  filter(origin == "JFK") %>% # 8,733
  filter(!is.na(wind_dir)) # 7,861 (90%)

hourly_weather_clean %>% 
  filter(origin == "JFK")

```


```{r}
# OlsonNames()
# 
# weather %>% 
#   summarise(max(time_hour))
# 
# weather_cleaned <- hourly_weather_clean %>% 
#   mutate(corrected_date_time = ymd_h(paste(year, month, day, hour, sep = "-")))
# 
# attr(weather_clean$time_hour, "tzone")
# attr(weather_cleaned$corrected_date_time, "tzone")
# 
# weather_cleaned %>% 
#   mutate(date_time = with_tz(ymd_h(paste(year, month, day, hour, sep = "-")),
#                              "America/New_York")) %>% 
#   head(10)
# 
# weather
# flights

```

# Annual Wind Data

Define maximum deviance between wind direction and closest angle of runway.
Sign of difference doesn't matter.

Wind magnitude creates a score = max_wind_dev * wind_speed

```{r}

wind_deviance_mag <- hourly_weather_clean %>% 
  group_by(time_hour, origin) %>% 
  mutate(min_wind_dev = if_else(origin == "EWR", 
                            min(abs(wind_dir - 40),
                                abs(wind_dir - (360 + 40)), # 40 degrees
                                abs(wind_dir - 220),
                                abs(wind_dir - 290), 
                                abs(wind_dir - 110)),
                            min(abs(wind_dir - 40),
                                abs(wind_dir - (360 + 40)), # 40 degrees
                                abs(wind_dir - 220),
                                abs(wind_dir - 310), 
                                abs(wind_dir - 130))),
         wind_magnitude = wind_speed * min_wind_dev) %>% 
  ungroup() %>% 
  select(time_hour, origin, wind_magnitude)

# wind_deviance_mag %>% 
#   group_by(wind_magnitude, origin) %>% 
#   summarise(avg_mag = mean(wind_magnitude)) %>% 
#   ggplot(aes(x = wind_dir, y = avg_mag, fill = origin)) +
#   geom_col() +
#   coord_polar() +
#   facet_wrap(~origin)

```

```{r}
flights
wind_deviance_mag

flights %>%
  filter(!is.na(is_delayed)) %>% 
  arrange(origin, time_hour, sched_dep_time) %>% 
  select(time_hour, sched_dep_time, origin, is_delayed) %>% 
  left_join(wind_deviance_mag, by = c("time_hour", "origin")) %>% 
  ggplot(aes(x = wind_magnitude, fill = is_delayed)) +
  geom_histogram(position = "fill")

```

# COMBINING FLIGHTS AND WEATHER DATA

```{r}

logregmodel_data <- flights %>% 
  filter(origin == "EWR") %>% 
  left_join(daily_weather_data, by = c("date", "origin")) %>%
  left_join(wind_deviance_mag, by = c("time_hour", "origin")) %>% 
  select(-time_hour, date) %>% 
  filter(!is_cancelled) %>% 
  select(-is_cancelled, -dest, -tailnum, -origin,-dep_delay) %>% 
  drop_na()
  


logregmodel_data

```
## Model to predict if flight will be delayed or not? (TRUE/FALSE)

- Logistic Regression Model

```{r, warning= FALSE, message=FALSE}

library(GGally)

logregmodel_data %>%
  ggpairs(progress = FALSE)

ggsave("raw_data/logreg_model.png",
      width = 20,
      height = 20)

```

```{r}

alias(is_delayed ~ ., data = logregmodel_data)

```


```{r}

logreg_model <- glm(is_delayed ~ .,
                    logregmodel_data,
                    family = binomial(link = "logit"))

logreg_model

```

AIC: 131300 (with wind magnitude)
AIC: 134000
AIC: 134100

```{r}

threshold <- 0.6

library(modelr)

logregmodel_with_pred <- logregmodel_data %>%
  add_predictions(logreg_model) %>% 
  select(is_delayed, pred) %>% 
  mutate(pred_thresh_0.6 = pred >= threshold)

conf_table <- logregmodel_with_pred %>% 
  tabyl(is_delayed, pred_thresh_0.6)

conf_table

```

Accuracy = (NTP + NTN) / N = (65417 + 7718) / (65417 + 2809 + 34976 + 7718)
                           = 65.9%
                           
```{r}

library(pROC)

```

```{r}

roc_obj_logreg <- logregmodel_with_pred %>% 
  roc(response = is_delayed, predictor = pred)

ggroc(data = roc_obj_logreg, legacy.axes = TRUE)

```

```{r}

auc(roc_obj_logreg)

```

