---
title: "R Notebook"
output: html_notebook
---

```{r}

library(tidyverse)
library(infer)

```

```{r}
# load in raw data set
flights <- read_csv("raw_data/flights.csv") 

prop_delay_flights <- flights %>% 
  group_by(origin) %>% 
  mutate(is_delayed = if_else(dep_delay >= 15, TRUE, FALSE),
         is_delayed = coalesce(is_delayed, FALSE))


jfk_prop_delays <- prop_delay_flights %>%
  filter(origin == "JFK") %>% 
  summarise(prop_delayed = mean(is_delayed)) %>% 
  pull()
  
lga_prop_delays <- prop_delay_flights %>%
  filter(origin == "LGA") %>% 
  summarise(prop_delayed = mean(is_delayed)) %>% 
  pull()

ewr_flights <- prop_delay_flights %>% 
  filter(origin == "EWR")

```

Null hypotheses (H0):         prop_delays EWR = prop_delays JFK
Alternative hypotheses (Ha):  prop_delays EWR != prop_delays JFK

```{r}

null_distribution_jfk <- ewr_flights %>%
  specify(response = is_delayed, success = "TRUE") %>% 
  hypothesize(null = "point", p = jfk_prop_delays) %>%
  generate(reps = 5000, type = "draw") %>%
  calculate(stat = "prop") #we specify that the stat is proportion (in the last lesson it was mean)

head(null_distribution_jfk)

```

```{r}

null_distribution_jfk %>%
  visualise(bins = 30)

```

```{r}

observed_stat <- ewr_flights %>% 
  specify(response = is_delayed, success = "TRUE") %>% 
  calculate(stat = "prop")

observed_stat

```

```{r}

null_distribution_jfk %>%
  visualise(bins = 30) + #remembering that it is a '+' here not '%>%' because using ggplot functionality to visualise
  shade_p_value(obs_stat = observed_stat, direction = "both") #it is a two-sided test as the alternative is whether the proportion is different in either left or left direction

```

```{r}

p_value_jfk <- null_distribution_jfk %>%
  get_p_value(obs_stat = observed_stat, direction = "both")

p_value_jfk

```

The proportion of flight delays at Newark airport is significantly different to the proportion of delays at JFK airport.

# LGA

```{r}

null_distribution_lga <- ewr_flights %>%
  specify(response = is_delayed, success = "TRUE") %>% 
  hypothesize(null = "point", p = lga_prop_delays) %>%
  generate(reps = 5000, type = "draw") %>%
  calculate(stat = "prop") #we specify that the stat is proportion (in the last lesson it was mean)

head(null_distribution_lga)

```

```{r}

null_distribution_lga %>%
  visualise(bins = 30)

```

```{r}

null_distribution_lga %>%
  visualise(bins = 30) + #remembering that it is a '+' here not '%>%' because using ggplot functionality to visualise
  shade_p_value(obs_stat = observed_stat, direction = "both") #it is a two-sided test as the alternative is whether the proportion is different in either left or left direction

```

```{r}

p_value_lga <- null_distribution_jfk %>%
  get_p_value(obs_stat = observed_stat, direction = "both")

p_value_lga

```

The proportion of flight delays at Newark airport is significantly different to the proportion of delays at LGA airport.

So why are the number of delays significantly different between EWR and JFK/LGA?