---
title: "R Notebook"
output: html_notebook
---

```{r}

library(tidyverse)
library(janitor)
library(lubridate)
library(modelr)
library(pROC)
library(glmulti)

```

```{r}

flights <- read_csv("clean_data/flights_clean.csv")

weather <- read_csv("clean_data/weather_clean.csv")

additional_weather <- read_csv("clean_data/additional_weather.csv")

planes <- read_csv("clean_data/planes_clean.csv")

```

# Model 1 - Simple Weather Model

a) weather only
b) weather and time
c) weather, time and carrier/plane

```{r}
# combine flights and planes data
flights_planes <- flights %>% 
  # left_join(weather, c("time_hour", "origin")) %>% 
  left_join(planes, "tailnum") %>%  
  filter(origin == "EWR") %>% 
  mutate(across(where(is.character), ~as_factor(.x))) %>% 
  drop_na()

model1_base_data <- flights_planes %>% 
  left_join(weather, c("time_hour", "origin")) %>%
  left_join(additional_weather, c("time_hour" = "datetime")) %>% 
  drop_na() %>% 
  select(-time_hour, -origin, -dest, -dep_delay, -wind_gust)

# Model 1a - weather only and is_delayed flag
model1a_data <- model1_base_data %>% 
  select(wind_dir:temperature, is_delayed)

model1a_data

# Model 1b - weather and time data
model1b_data <- model1_base_data %>% 
  select(date:hour, wind_dir:temperature, is_delayed)

model1b_data

# Model 1c - weather, time and plane data
model1c_data <- model1_base_data %>% 
  select(-tailnum, -seats)

model1c_data

```

```{r}

model1a <- glm(is_delayed ~ . - temperature,
              data = model1a_data,
              family = binomial(link = "logit"))

summary(model1a)

model1a_pred <- model1a_data %>% 
  add_predictions(model1a, type = "response")

roc_obj_model1a <- model1a_pred %>% 
  roc(response = is_delayed, predictor = pred)

ggroc(roc_obj_model1a,
      legacy.axes = TRUE)

auc(roc_obj_model1a)

```

```{r}

model1b <- glm(is_delayed ~ .,
              data = model1b_data,
              family = binomial(link = "logit"))

summary(model1b)

model1b_pred <- model1b_data %>% 
  add_predictions(model1b, type = "response")

roc_obj_model1b <- model1b_pred %>% 
  roc(response = is_delayed, predictor = pred)

ggroc(roc_obj_model1b,
      legacy.axes = TRUE)

auc(roc_obj_model1b)

```

```{r}

model1c <- glm(is_delayed ~ .,
              data = model1c_data,
              family = binomial(link = "logit"))

summary(model1c)

model1c_pred <- model1c_data %>% 
  add_predictions(model1c, type = "response")

roc_obj_model1c <- model1c_pred %>% 
  roc(response = is_delayed, predictor = pred)

ggroc(roc_obj_model1c,
      legacy.axes = TRUE)

auc(roc_obj_model1c)

```

## Analysis of Variation - Model 1c

```{r}

model1c_no_manufacturer <- glm(is_delayed ~ . - manufacturer,
                               data = model1c_data,
                               family = binomial(link = "logit"))

anova(model1c, model1c_no_manufacturer)

model1c_no_size_cat <- glm(is_delayed ~ . - size_cat,
                               data = model1c_data,
                               family = binomial(link = "logit"))

anova(model1c, model1c_no_size_cat)

model1c_no_carrier <- glm(is_delayed ~ . - carrier,
                               data = model1c_data,
                               family = binomial(link = "logit"))

anova(model1c, model1c_no_carrier)

```
# Model 1d - glmulti output - no interactions (doesn't work)

```{r}

model1d <- glmulti(is_delayed ~ .,
                   data = model1c_data,
                   level = 2,               # Interactions considered
                   method = "g",            # Genetic algorithm approach
                   crit = "aic",            # AIC as criteria
                   confsetsize = 10,        # Keep 10 best models
                   marginality = TRUE,      # consider pairs only if both main effects in model
                   plotty = F, 
                   report = T, # No plots, but provide interim reports
                   fitfunction = "glm",     # glm function
                   family = binomial(link = "logit")) # binomial family for logistic regression

summary(model1d)

model1d_pred <- model1c_data %>% 
  add_predictions(model1c, type = "response")

roc_obj_model1d <- model1d_pred %>% 
  roc(response = is_delayed, predictor = pred)

ggroc(roc_obj_model1d,
      legacy.axes = TRUE)

auc(roc_obj_model1d)

```

```{r}

ggroc(list("model1a" = roc_obj_model1a,
           "model1b" = roc_obj_model1b,
           "model1c" = roc_obj_model1c,
           "model1d" = roc_obj_model1d),
      legacy.axes = TRUE)

auc(roc_obj_model1a)
auc(roc_obj_model1b)
auc(roc_obj_model1c)
auc(roc_obj_model1d)

```
 
# Model 1d - model 1c using glmulti




```{r}

model1d <- glmulti(
  is_delayed~1+carrier+date+month+hour+age+wind_speed+visib+humidity+pressure+temperature,
  data = model1c_data,
  level = 2,               # Interactions considered
  method = "g",            # Genetic algorithm approach
  crit = "bic",            # BIC as criteria
  confsetsize = 10,        # Keep 10 best models
  marginality = TRUE,      # consider pairs only if both main effects in model
  plotty = F, 
  report = T,              # No plots, but provide interim reports
  fitfunction = "glm",     # glm function
  family = binomial(link = "logit")) # binomial family for logistic regression

```

After 14750 models:
Best model: is_delayed~1+carrier+date+month+hour+age+wind_speed+visib+humidity+pressure+temperature
Crit= 101072.398497548
Mean crit= 101085.359246972


# Model 2 - Add daily weather data

```{r}

daily_weather <- read_csv("clean_data/daily_weather.csv")

daily_weather

model2_data <- ewr_flights_weather %>% 
  left_join(daily_weather, by = c("date", "origin")) %>% 
  mutate(across(where(is.character), ~as_factor(.x))) %>% 
  select(-origin, -dest, -tailnum, -dep_delay, -time_hour, -wind_gust,
         -carrier, -wind_dir)

```

```{r}

model2 <- glm(is_delayed ~ .,
              data = model2_data,
              family = binomial(link = "logit"))

summary(model2)

```

```{r}

model2_pred <- model2_data %>% 
  add_predictions(model2, type = "response")

roc_obj_model2 <- model2_pred %>% 
  roc(response = is_delayed, predictor = pred)

ggroc(list("model1" = roc_obj_model1,
           "model2" = roc_obj_model2),
      legacy.axes = TRUE)

auc(roc_obj_model1)
auc(roc_obj_model2)


```

# Model 3 - Using additional hourly weather data (humidity, pressure, temp)
Only covers the period from Jan 2017 to Nov 2017

```{r}

additional_weather <- read_csv("clean_data/additional_weather.csv")

model3_data <- ewr_flights_weather %>% 
  left_join(daily_weather, by = c("date", "origin")) %>% 
  left_join(additional_weather, c("time_hour" = "datetime")) %>% 
  mutate(across(where(is.character), ~as_factor(.x))) %>% 
  select(-origin, -dest, -tailnum, -dep_delay, -time_hour, -wind_gust,
         -carrier, -(tavg:tmin), -wind_dir, -humidity) %>% 
  drop_na()

```

```{r}

model3 <- glm(is_delayed ~ .,
              data = model3_data,
              family = binomial(link = "logit"))

summary(model3)

```

```{r}

model3_pred <- model3_data %>% 
  add_predictions(model3, type = "response")

roc_obj_model3 <- model3_pred %>% 
  roc(response = is_delayed, predictor = pred)

ggroc(list("model1" = roc_obj_model1,
           "model2" = roc_obj_model2,
           "model3" = roc_obj_model3),
      legacy.axes = TRUE)

auc(roc_obj_model1)
auc(roc_obj_model2)
auc(roc_obj_model3)

```
