---
title: "R Notebook"
output: html_notebook
---

```{r}

library(tidyverse)
library(janitor)
library(lubridate)
library(modelr)
library(pROC)
library(glmulti)
library(fastDummies)

```

```{r}

flights <- read_csv("clean_data/flights_clean.csv")

weather <- read_csv("clean_data/weather_clean.csv")

additional_weather <- read_csv("clean_data/additional_weather.csv")

planes <- read_csv("clean_data/planes_clean.csv")

daily_weather <- read_csv("clean_data/daily_weather.csv") %>% 
  select(-(tavg:tmin))

```

# Model 1 - Simple Weather Model

a) weather only
b) weather and time
c) weather, time and carrier/plane

```{r}
# combine flights and planes data
flights_planes <- flights %>% 
  left_join(planes, "tailnum") %>%  
  filter(origin == "EWR") %>% 
  mutate(across(where(is.character), ~as_factor(.x))) %>%
  drop_na()


model1_base_data <- flights_planes %>% 
  left_join(weather, c("time_hour", "origin")) %>%
  left_join(additional_weather, c("time_hour" = "datetime")) %>%
  left_join(daily_weather, c("date", "origin")) %>% 
  drop_na() %>% 
  select(-time_hour, -origin, -dest, -dep_delay, -wind_gust, -tailnum, -seats,
         -manufacturer, -wind_dir, -date)

  
# model1_base_data_dummies <- model1_base_data %>% 
#     dummy_cols(select_columns = "manufacturer",
#              remove_first_dummy = TRUE,
#              remove_selected_columns = TRUE) %>% 
#   dummy_cols(select_columns = "carrier",
#              remove_first_dummy = TRUE,
#              remove_selected_columns = TRUE)
  
```

## create test/train split

```{r}

n_data <- nrow(model1_base_data)
test_index <- sample(1:n_data, size = n_data * 0.2)

test  <- slice(model1_base_data, test_index)
train <- slice(model1_base_data, -test_index)

train

```



```{r}
# Model 1a - weather only and is_delayed flag
model1a_data <- train %>% 
  select(wind_speed:has_weather_warning, is_delayed)

model1a_data

# Model 1b - weather and time data
model1b_data <- train %>% 
  select(wday:hour, wind_speed:has_weather_warning, is_delayed)

model1b_data

# Model 1c - weather, time and plane/carrier data
model1c_data <- train

model1c_data

```

```{r}

model1a <- glm(is_delayed ~ .,
              data = model1a_data,
              family = binomial(link = "logit"))

summary(model1a)

model1a_pred <- model1a_data %>% 
  add_predictions(model1a, type = "response")

roc_obj_model1a <- model1a_pred %>% 
  roc(response = is_delayed, predictor = pred)

ggroc(roc_obj_model1a,
      legacy.axes = TRUE)

auc(roc_obj_model1a)

```

```{r,eval=FALSE}

model1a_glmulti_data <- model1a_data

# model 1a using glmulti genetic algorithm
model1a_glmulti <- glmulti(
  is_delayed ~ ., 
  data = model1a_glmulti_data,
  level = 1,               # Interactions considered
  method = "h",            # Exhaustive approach
  crit = "aic",            # BIC as criteria
  confsetsize = 1,        # Keep 10 best models
  marginality = TRUE,      # consider pairs only if both main effects in model
  #minsize = 12,             # minsize, maxsize and marginality here force 
  #maxsize = 12,             # inclusion of a single pair beyond the five main effects
  plotty = F, 
  report = T,              # No plots, but provide interim reports
  fitfunction = "glm",     # glm function
  family = binomial(link = "logit")) # binomial family for logistic regression

summary(model1a_glmulti)

# best model: is_delayed ~ 1 + wind_speed + visib + min_wind_dev + wind_magnitude + humidity + pressure + temperature + prcp + snow + snow_depth + has_weather_warning 

```

```{r,eval=FALSE}

model1a_glmulti_data <- model1a_data

# model 1a using glmulti genetic algorithm
model1a_glmulti_interactions <- glmulti(
  is_delayed ~ 1 + wind_speed + visib + min_wind_dev + wind_magnitude + humidity
  + pressure + temperature + prcp + snow + snow_depth + has_weather_warning, 
  data = model1a_glmulti_data,
  level = 2,               # Interactions considered
  method = "h",            # Exhaustive approach
  crit = "bic",            # BIC as criteria
  confsetsize = 1,        # Keep 10 best models
  marginality = TRUE,      # consider pairs only if both main effects in model
  minsize = 12,             # minsize, maxsize and marginality here force 
  #maxsize = 12,             # inclusion of a single pair beyond the five main effects
  plotty = F, 
  report = T,              # No plots, but provide interim reports
  fitfunction = "glm",     # glm function
  family = binomial(link = "logit")) # binomial family for logistic regression

summary(model1a_glmulti_interactions)

```


Delay threshold >= 0: 0.6463
Delay threshold >= 10: 0.6598
Delay threshold >= 20: 0.6687

```{r}

model1b <- glm(is_delayed ~ .,
              data = model1b_data,
              family = binomial(link = "logit"))

summary(model1b)

model1b_pred <- model1b_data %>% 
  add_predictions(model1b, type = "response")

roc_obj_model1b <- model1b_pred %>% 
  roc(response = is_delayed, predictor = pred)

ggroc(roc_obj_model1b,
      legacy.axes = TRUE)

auc(roc_obj_model1b)

```

Delay threshold >= 0: 0.7191
Delay threshold >= 10: 0.7377
Delay threshold >= 20: 0.7495



```{r}

model1c <- glm(is_delayed ~ .,
              data = model1c_data,
              family = binomial(link = "logit"))

summary(model1c)

model1c_pred <- model1c_data %>% 
  add_predictions(model1c, type = "response")

roc_obj_model1c <- model1c_pred %>% 
  roc(response = is_delayed, predictor = pred)

ggroc(roc_obj_model1c,
      legacy.axes = TRUE)

auc(roc_obj_model1c)

```

Delay threshold >= 0: 0.7292
Delay threshold >= 10: 0.7453
Delay threshold >= 20: 0.757

With engine: Area under the curve: 0.7292
Without engine: Area under the curve: 0.729
Without carrier: Area under the curve: 0.7214

## Analysis of Variation - Model 1c

```{r}

# model1c_no_manufacturer <- glm(is_delayed ~ . - manufacturer,
#                                data = model1c_data,
#                                family = binomial(link = "logit"))
# 
# anova(model1c, model1c_no_manufacturer)

model1c_no_engine <- glm(is_delayed ~ . - engine,
                               data = model1c_data,
                               family = binomial(link = "logit"))

anova(model1c, model1c_no_engine)

model1c_no_carrier <- glm(is_delayed ~ . - carrier,
                               data = model1c_data,
                               family = binomial(link = "logit"))

anova(model1c, model1c_no_carrier)

```


```{r}

ggroc(list("model1a" = roc_obj_model1a,
           "model1b" = roc_obj_model1b,
           "model1c" = roc_obj_model1c),
      legacy.axes = TRUE)

auc(roc_obj_model1a)
auc(roc_obj_model1b)
auc(roc_obj_model1c)

```
Delay threshold >= 0:
Area under the curve: 0.6463
Area under the curve: 0.7191
Area under the curve: 0.7292

Delay threshold >= 10:
Area under the curve: 0.6598
Area under the curve: 0.7377
Area under the curve: 0.7453

Delay threshold >= 20:
Area under the curve: 0.6687
Area under the curve: 0.7495
Area under the curve: 0.757


# Cross-Validation

```{r, eval=FALSE}

library(caret)

model1_cv_data <- model1c_data %>% 
  mutate(is_delayed = as.factor(if_else(is_delayed, "t", "f")),
         engine = make.names(engine)) %>% 
  select(-wind_magnitude)

train_control <- trainControl(method = "repeatedcv", 
                              number = 5,
                              repeats = 100,
                              savePredictions = TRUE, 
                              classProbs = TRUE, 
                              summaryFunction = twoClassSummary)

model <- train(is_delayed ~ . - humidity - min_wind_dev,
               data = model1_cv_data,
               trControl = train_control,
               method = "glm",
               family = binomial(link = 'logit'))

summary(model)

```

Delay threshold >= 0: 79314
Delay threshold >= 10: 70090
Delay threshold >= 20: 61797

```{r, eval=FALSE}

model

```



```{r}

model$result

```

AUC
Delay threshold >= 0: 0.7285887
Delay threshold >= 10: 0.7445493
Delay threshold >= 20: 0.7562068

```{r, eval=FALSE}

model_cv_data_pred <- test %>%
  mutate(is_delayed = as.factor(if_else(is_delayed, "t", "f")),
         engine = make.names(engine)) %>% 
  add_predictions(model, var = "response") %>% 
  select(is_delayed, response)

model_cv_data_pred %>% 
  summarise(accuracy = sum(is_delayed == response) / n())

```

Delay_threshold >= 0: 68.6%
Delay_threshold >= 10: 74.4%
Delay_threshold >= 20: 78.9%

77% accuracy against test data set. (0.7699677)
Adding wday, removing date: 77.1% accuracy (0.7708658)	


```{r}
model2_base_data <- flights_planes %>% 
  left_join(weather, c("time_hour", "origin")) %>%
  left_join(additional_weather, c("time_hour" = "datetime")) %>%
  left_join(daily_weather, c("date", "origin")) %>% 
  drop_na() %>% 
  select(-time_hour, -origin, -dest, -dep_delay, -wind_gust, -tailnum, -seats,
         -manufacturer, -date, -wind_magnitude)


model2_cv_data <- model2_base_data %>% 
  mutate(is_delayed = as.factor(if_else(is_delayed, "t", "f")),
         engine = make.names(engine))

model2 <- glm(is_delayed ~ .,
              model2_base_data,
              family = binomial(link = "logit"))

summary(model2)

model2a <- glm(is_delayed ~ . + wind_dir:wind_speed - wind_dir,
              model2_base_data,
              family = binomial(link = "logit"))

summary(model2a)


```



```{r,eval=FALSE}
train_control <- trainControl(method = "repeatedcv", 
                              number = 5,
                              repeats = 100,
                              savePredictions = TRUE, 
                              classProbs = TRUE, 
                              summaryFunction = twoClassSummary)

model <- train(is_delayed ~ . - humidity - min_wind_dev,
               data = model1_cv_data,
               trControl = train_control,
               method = "glm",
               family = binomial(link = 'logit'))

summary(model)

```

# glmulti - all data model

```{r, message = FALSE, error= FALSE, warning = FALSE,eval=FALSE}

library(GGally)

model1c_data %>% 
  select(is_delayed, 1:10) %>% 
  ggpairs(progress= FALSE)

```

```{r, message = FALSE, error= FALSE, warning = FALSE,eval=FALSE}
model1c_data %>% 
  select(is_delayed, 11:19) %>% 
  ggpairs(progress= FALSE)

```


```{r}
model3_glmulti_data <- model1c_data %>% 
  mutate(is_delayed = as.numeric(is_delayed)) %>% 
  select(-wday, -engine, -carrier)

model2_glmulti_data <- model1c_data %>% 
  mutate(is_delayed = as.numeric(is_delayed))

```


```{r,eval=FALSE}
# model 1a using glmulti genetic algorithm
model1c_glmulti_interactions <- glmulti(
  is_delayed ~ .,
  data = model1c_glmulti_data,
  level = 2,               # Interactions considered
  method = "g",            # Exhaustive approach
  crit = "bic",            # BIC as criteria
  confsetsize = 10,        # Keep 10 best models
  marginality = TRUE,      # consider pairs only if both main effects in model
  #minsize = 18,             # minsize, maxsize and marginality here force 
  #maxsize = 12,             # inclusion of a single pair beyond the five main effects
  plotty = F, 
  report = T,              # No plots, but provide interim reports
  fitfunction = "glm",     # glm function
  family = binomial(link = "logit")) # binomial family for logistic regression

summary(model1c_glmulti_interactions)


```

Best model: is_delayed~1+size_cat+month+hour+plane_age+wind_speed+visib+min_wind_dev+wind_magnitude+humidity+pressure+temperature+prcp+snow+snow_depth+has_weather_warning+wind_speed:hour+wind_speed:plane_age+visib:month+wind_magnitude:wind_speed+pressure:month+pressure:hour+pressure:plane_age+pressure:humidity+temperature:month+temperature:visib+temperature:min_wind_dev+temperature:pressure+prcp:month+prcp:plane_age+prcp:min_wind_dev+prcp:wind_magnitude+prcp:pressure+prcp:temperature+snow:hour+snow:wind_speed+snow:min_wind_dev+snow:temperature+snow_depth:hour+snow_depth:pressure+snow_depth:temperature+has_weather_warning:month+has_weather_warning:hour+has_weather_warning:wind_speed+has_weather_warning:min_wind_dev+has_weather_warning:humidity+has_weather_warning:snow+has_weather_warning:snow_depth+size_cat:month+size_cat:hour+size_cat:plane_age+size_cat:has_weather_warning

# Test out best 3 models from glmulti
model1c = best model without interactions
glmulti_model1 = glmulti (min 12 terms - exhaustive)
glmulti_model2 = glmulti model (exhaustive)
glmulti_model3 = glmulti model (genetic algorithm) - removed chr variables

```{r}
logregmodel1 <- glm(is_delayed ~ .,
                    data = model1c_data,
                    family = binomial(link = "logit"))

glmulti_model1 <- glm(is_delayed~1+month+hour+plane_age+wind_speed+visib+wind_magnitude+hour:month+plane_age:month+wind_speed:month+wind_speed:hour+visib:month+visib:hour+visib:plane_age+visib:wind_speed+wind_magnitude:month+wind_magnitude:hour+wind_magnitude:wind_speed+wind_magnitude:visib, 
                      data = model2_glmulti_data,
                      family = binomial(link = "logit"))

glmulti_model2 <- glm(is_delayed~1+month+hour+wind_speed+visib+min_wind_dev+visib:month+visib:hour+min_wind_dev:wind_speed,
                      data = model2_glmulti_data,
                      family = binomial(link = "logit"))

glmulti_model3 <- glm(is_delayed~1+size_cat+month+hour+plane_age+wind_speed+visib+min_wind_dev+wind_magnitude+humidity+pressure+temperature+prcp+snow+snow_depth+has_weather_warning+wind_speed:hour+wind_magnitude:month+humidity:month+temperature:wind_speed+temperature:visib+temperature:pressure+prcp:month+prcp:hour+prcp:min_wind_dev+prcp:pressure+snow:hour+snow:temperature+snow_depth:humidity+snow_depth:temperature+has_weather_warning:month+has_weather_warning:hour+has_weather_warning:wind_speed+has_weather_warning:humidity+size_cat:hour,
                      data = model3_glmulti_data,
                      family = binomial(link = "logit"))
```


```{r}
# base model
logregmodel1_pred <- model1c_data %>%
  add_predictions(logregmodel1, type = "response")

roc_obj_logregmodel1 <- logregmodel1_pred %>% 
  roc(response = is_delayed, predictor = pred)

#glmulti model 1
glmulti_model1_pred <- model2_glmulti_data %>%
  add_predictions(glmulti_model1, type = "response")

roc_obj_glmulti_model1 <- glmulti_model1_pred %>% 
  roc(response = is_delayed, predictor = pred)

#glmulti model 2
glmulti_model2_pred <- model2_glmulti_data %>%
  add_predictions(glmulti_model2, type = "response")

roc_obj_glmulti_model2 <- glmulti_model2_pred %>% 
  roc(response = is_delayed, predictor = pred)

#glmulti model 3
glmulti_model3_pred <- model3_glmulti_data %>%
  add_predictions(glmulti_model3, type = "response")

roc_obj_glmulti_model3 <- glmulti_model3_pred %>% 
  roc(response = is_delayed, predictor = pred)

# accuracies vs test data set

true_thresh <- 0.5

logregmodel1_pred %>% 
  mutate(pred = if_else(pred >= true_thresh, TRUE, FALSE)) %>% 
  summarise(accuracy = sum(is_delayed == pred) / n())

glmulti_model1_pred %>% 
  mutate(pred = if_else(pred >= true_thresh, TRUE, FALSE)) %>% 
  summarise(accuracy = sum(is_delayed == pred) / n())
  
glmulti_model2_pred %>% 
  mutate(pred = if_else(pred >= true_thresh, TRUE, FALSE)) %>% 
  summarise(accuracy = sum(is_delayed == pred) / n())

glmulti_model3_pred %>% 
  mutate(pred = if_else(pred >= true_thresh, TRUE, FALSE)) %>% 
  summarise(accuracy = sum(is_delayed == pred) / n())

model1c_data


```


```{r}

ggroc(list("model1c" = roc_obj_logregmodel1,
           "glmulti1" = roc_obj_glmulti_model1,
           "glmulti2" = roc_obj_glmulti_model2,
           "glmulti3" = roc_obj_glmulti_model3),
      legacy.axes = TRUE)

auc(roc_obj_logregmodel1)
auc(roc_obj_glmulti_model1)
auc(roc_obj_glmulti_model2)
auc(roc_obj_glmulti_model3)

```



```{r, eval = FALSE}

model_cv_data_pred %>% 
  summarise(accuracy = sum(is_delayed == response) / n())


```

# Random Forest Model

```{r}

library(ranger)

rf_data <- model1c_data %>% 
  mutate(is_delayed = as.factor(is_delayed)) %>% 
  select(-wind_magnitude)

rf_classifier <- ranger(is_delayed ~ .,
                        data = rf_data,
                        importance = "impurity",
                        num.trees = 5000,
                        mtry = 2,
                        min.node.size = 5)

rf_classifier
```

```{r}

sort(importance(rf_classifier), decreasing = TRUE)

```

```{r}

delays_pred <- rf_data %>%
  mutate(pred = predict(rf_classifier, data = model1c_data)$predictions)

```

```{r}

library(caret)

confusionMatrix(delays_pred$pred, delays_pred$is_delayed)

```

```{r}

delays_pred_test_data <- test %>%
  mutate(is_delayed = as.factor(is_delayed)) %>% 
  mutate(pred = predict(rf_classifier, data = test)$predictions)

confusionMatrix(delays_pred_test_data$pred, delays_pred_test_data$is_delayed)

```

